A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-21 manuscript/manuscript.tex
ms updates

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_7.pdf}

@@

Finally, we entertain two more speculative predictions. First, the equilibrium
additive genic fitness variance. Our PhastCons Priority model DFE estimates and
negative selection model imply a $\widehat{V}_a = 0.45$ for the default grid,
and $\widehat{V}_a = 0.040$ for the constrained grid. This is a considerable
level of uncertainty---over an order of magnitude---due to a difference in the
selection grid. There are no other known direct estimates in humans of this
quantity for comparison. When $2N \gg 1$, theory suggests $V_a \approx
U_D\bar{s}$ where $U_D$ is the diploid rate of strongly deleterious mutations.
Thus, the uncertainty is a byproduct of the different average estimates of
$\bar{s}$ across the alternate grids. Since both this work and
\textcite{Murphy2022-sj} find the single-feature CADD 6\% track fits well, we
can use this to gain intuition through some rough calculations. In this case,
$U_D \approx 6\% \times 2 \times 3 \times 10^{9} \mu \approx 3.6 \times 10^{8}
\mu$. For $\mu \approx 1.5 \times 10^{-8}$, $U_D \approx 5.2$. Under the
constrained grid, ${\bar{s}} \approx 0.0079$ and thus $V_a \approx 0.043$; with
the default grid $\bar{s} \approx 0.043$, and $V_a \approx 0.23$. Often fitness
reductions of $\approx 1\%$ are assumed for the most highly-conserved regions
(XXX, Keightly, Lynch); for comparison, this leads to $V_a \approx 0.054$.

XXX HERE  USE CADD PREDICTIONS

By theory, this relative fitness (scaled by the mean squared) is
bounded from above by the Index of Selection, defined as the theoretic relative
fitness variance if fitness were perfectly heritable \parencite{Crow1958-pc}.
Based on rough calculation by Crow, this could range between 0.982-3.286 in
humans. However, this is based on modern-day mortality figures that may not
extrapolate to the timescales that determine genome-wide patterns of diversity.

Second, we can calculate the reduction in relative fitness per generation due
to deleterious substitutions. Since \textcite{Muller1950-ta}, there has been
considerable interest in whether mildly deleterious mutations, perhaps due to
the relaxation of selection pressures, could accumulate and fix at a faster
pace in humans.

Here, the selection grid choice has little impact
on the estimates: we calculate that across the genome, the loss in relative
fitness per generation is 0.00160\% for the default grid, and 0.00119\% for the
constrained grid.

@@

Given that our simulations indicated that $B(x)$ is more accurately predicted
when we use local-rescaling, we modified our composite-likelihood method to be
run a second time. Intuitively, this is based on the notion that if a neutral
allele experiences an effective population size of $B(x)N$, so too should a
selected allele, and this should be considered in how the SC16 equations are
solved. Immediately, there is an important caveats to note about this analysis.
Local-rescaling is done using the first round of maximum likelihood estimates,
and then re-solving the system of equations with a population size of $B(x)N$.
This could introduce the possibility of some circularity, since the $B(x)$ at
this position includes the contribution of the focal segment that is being
rescaled. However, the qualitative results do not change based on whether the
$B(x)$ is at averaged at the 1 kbp level or at the megabase scale; for the
latter, a single segment makes little contribution to the average.

With this caveat noted, there are three important but tentative results to draw
from this analysis. First, the estimated mutation rate is implausibly higher,
reaching the upper limit of our optimization boundary of $\mu = 8 \times
10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
Material XXX). Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
are about one percentage point higher than the unrescaled model. This improved
fit could be an artifact of the circularity, so it is unimportant. Third, the
DFE estimates for PhastCons regions is now U-shaped, with $75-77\%$ of mass
being placed on a weakly deleterious class, $s=10{^-5}$. Interestingly, this is
the first of all of our models where such an appreciable mass has been placed
on a midpoint in our grid; in all other cases, non-strongly estimates were
neutral ($s=10^{-8}$). Finally, predicted pairwise diversity is nearly
identical to our original, non-locally-rescaled fit (see Supplementary
Materials XXX). This

Overall, this leads to an open end in our work. If selected sites along the
genome were to experience a draft effective population size that is closer to
our locally-rescaled version, then these quantitative genetic models of linked
selection

the fit to the linked selection signal is very close under
different parameter combinations (Supplementary Materials XXX). The fits with
local-rescaling

Intuitively, this is based on the notion that if a neutral allele experiences
an effective population size of $B(x)N$, so too should a selected allele. We
find in simulations this reduces error around $2 N_e s \approx 1$.
Unfortunately, employing this in our composite-likelihood method would create
some circularity, since the local $B(x)$ experienced at a conserved segment
includes the reduction due to the segment itself. There are not computationally
feasible ways to avoid this issue in the current implementation.

>>>>

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_7_mod.pdf}

@@

@@

Given the observed DFE-substitution rate contradiction and that our simulations
indicated that $B(x)$ is more accurately predicted when we use local-rescaling,
we modified our composite-likelihood method to so that it can be run a second
time, on B' maps locally-rescaled by the predicted $\widehat{B}(x)$ from the
initial fit. Intuitively, this is based on the notion that if a neutral allele
experiences an effective population size of $B(x)N$, so too should a selected
allele, and this should be considered in how the SC16 equations are solved.
There is an important caveat about this analysis. Since local-rescaling is done
using the first round of maximum likelihood estimates, there is a possibility
of some circularity, since the $B(x)$ at this position includes the
contribution of the focal segment that is being rescaled. However, the results
do not change based on whether the $B(x)$ is at averaged at the 1 kbp level or
at the megabase scale; for the latter, a single segment makes little
contribution to the average.

With this caveat noted, there are five important but tentative results to draw
from this analysis. First, the estimated mutation rate is implausibly higher,
reaching the upper limit of our optimization boundary of $\mu = 8 \times
10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
Material XXX). When we re-ran optimization with an upper limit of $\mu = 1
\times 10^{-7}$, the mutation rate was again estimated to this (implausibly
large) upper limit, with no considerable changes to model fit or DFE estimates.
Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$ are about one
percentage point higher than the unrescaled model. This improved fit could be
an artifact of the circularity, so it is unimportant. Third, the DFE estimates
for PhastCons regions is now U-shaped (Supplementary Materials Figure
\ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
deleterious class, $s=10{^-5}$. Interestingly, this is the first of all of our
models where such an appreciable mass has been placed on a midpoint in our
grid; in all other cases, non-strongly estimates were neutral ($s=10^{-8}$).
Fourth, predicted pairwise diversity is nearly identical to our original,
non-locally-rescaled fit (see Supplementary Materials Figure
\ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of
fixation. We find that local-rescaling the PhastCons Priority, leads the
predicted substitution rates in PhastCons regions to be much closer to observed
levels (blue line, Figure \ref{fig:figure-6}).

Finally, as a robustness check, we investigated the extent to which
local-rescaling differed from the B' maps across selection parameters. We find
minor differences between the locally-rescaled and standard B' maps in the $0.2
\le 2N_e s \le 2$ range; the maps were identical under strong selection ($2N_e
s= 20$) as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}).
Overall, this suggests that models that fit the linked selection are worryingly
sensitive the theoretic B' values in the $2Ns \le 1$ domain. Local-rescaling
turns out to only slightly alter the B' maps, yet significantly modifies the
DFE estimates. This brings the deleterious substitution rate in agreement with
observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$;
however, the maximum likelihood estimate of mutation is implausibly high.
